---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- use devtools::build_readme() to update -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# EAVA

Verbal Autopsy Expert Algorithm is a computer-coded verbal autopsy algorithm (CCVA) which assigns causes of death using responses obtained from the 2016 World Health Organization Verbal Autopsy Questionnaire. Historically, causes of death have been assigned by physician coding of questionnaire responses. As CCVAs become more widely used, physician time can be reallocated to other priorities in low resources settings.

## Installation

You can install the development version of EAVA from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("emilybrownwilson/EAVA")
```

## Example

This is a basic example which shows you how to use the package:

```{r, warning = FALSE, message = FALSE, results = 'hide'}
# load EAVA package
library(EAVA)

# load embedded example data
data <- as.data.frame(data_public)

# run odk2EAVA() using example data
output <- odk2EAVA(data, id_col  = "comsa_id")

# run codEAVA for neonates and children 1-to-59 months of age
EAVA_neonate <- codEAVA(output, "neonate")
EAVA_child <- codEAVA(output, "child")
```

```{r, warning = FALSE, message = FALSE}
# check results
head(EAVA_neonate)
head(EAVA_child)
```
