---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- use devtools::build_readme() to update -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# EAVA

Verbal Autopsy Expert Algorithm (EAVA) is a computer-coded verbal autopsy algorithm (CCVA) which evaluates signs and symptoms present at the time of death and reported in the 2016 WHO Verbal Autopsy Questionnaire. If a decedent meets the diagnostic criteria for more than one cause, a single cause of death is determined using an age-specific hierarchy of causes. Historically, causes of death have been assigned by physician coding of questionnaire responses. EAVA produces timely, reproducible cause of death assignments, enabling physicians to address other priorities in low resource settings.

## Installation

You can install the development version of EAVA from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("emilybrownwilson/EAVA")
```

Or load the package after installing from CRAN:
```{r}
# load EAVA package
#install.packages("EAVA")
library(EAVA)
```

# Codebook

ccva_codebook maps 2016 WHO VA questionnaire questions and their corresponding names to openVA variables. 

```{r, warning = FALSE, message = FALSE}
# expolore symtptoms and function mapping
data(ccva_codebook)

# View first few mappings
head(ccva_codebook)

# Look up a specific openVA variable
subset(ccva_codebook, openVA_var == "i147o")

# Trace a WHO variable to openVA variables
subset(ccva_codebook, who2016_var == "id10147")
```

## EAVA Example

This is a basic example which shows you how to load the EAVA package and run the functions. odk2EAVA() converts 2016 WHO VA questionnaire variable names and responses to formatting conventions used for openVA. codEAVA() assign CoDs by EAVA for neonates and children 1-59 months of age.

```{r, warning = FALSE, message = FALSE, results = 'hide'}
# load EAVA package
library(EAVA)

# load embedded example data
data <- as.data.frame(data_public)

# run odk2EAVA() using example data
output <- odk2EAVA(data, id_col  = "comsa_id")

# run codEAVA for neonates and children 1-to-59 months of age
EAVA_neonate <- codEAVA(output, "neonate")
EAVA_child <- codEAVA(output, "child")
```

```{r, warning = FALSE, message = FALSE}
# check results
head(EAVA_neonate)
head(EAVA_child)
```




