---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- use devtools::build_readme() to update -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# EAVA

Verbal Autopsy Expert Algorithm is a computer-coded verbal autopsy algorithm (CCVA) which assigns causes of death using responses obtained from the 2016 World Health Organization Verbal Autopsy Questionnaire. Historically, causes of death have been assigned by physician coding of questionnaire responses. CCVAs offer several advantages, providing reliable, reproducible cause of death determination and as well as enabling physicians to address other priorities in low resources settings.

## Installation

You can install the development version of EAVA from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("emilybrownwilson/EAVA")
```

## Example

This is a basic example which shows you how to load the packag and run the functions:

```{r, warning = FALSE, message = FALSE, results = 'hide'}
# load EAVA package
library(EAVA)

# load embedded example data
data <- as.data.frame(data_public)

# run odk2EAVA() using example data
output <- odk2EAVA(data, id_col  = "comsa_id")

# run codEAVA for neonates and children 1-to-59 months of age
EAVA_neonate <- codEAVA(output, "neonate")
EAVA_child <- codEAVA(output, "child")
```

```{r, warning = FALSE, message = FALSE}
# check results
head(EAVA_neonate)
head(EAVA_child)
```

# Codebook

The signs and symptoms which EAVA uses to determine causes of death are presented in the codebook

```{r, warning = FALSE, message = FALSE}
# expolore symtptoms and function mapping
data(ccva_codebook)

# View first few mappings
head(ccva_codebook)

# Look up a specific openVA variable
subset(ccva_codebook, openVA_var == "i147o")

# Trace a WHO variable to openVA variables
subset(ccva_codebook, who2016_var == "Id10147")
```


